### 2020-12-21 latex/acrobot.tex
Finished the acrobot experiments section

\section{The Acrobot Constraint}
One may be tempted to design a constraint of the form \(q_a = \bar{q}_a\sin(\theta)\) 
with \(\theta := \arctan_2(p_u,q_u)\), since a similar approach 
was so effective for the VLP in Chapter \ref{ch:vlp}.
Unfortunately, this constraint is not regular, and it is difficult to find any
VNHC of the form \(q_a = f(\theta)\) where regularity can be proven easily.
Instead, we will develop a constraint \(h(q,p) = q_a - f(p_u)\) because these
constraints are always regular (as per Proposition \ref{prop:acrobot-fpu-regular}).

@@

By the same arguments presented in this section, the Poincar\'{e} section
satisfies \( P_\mathcal{O}(\mu_0) \leq \mu_0 - \gamma\) when \(I < 0\). 
In this case, the constraint dissipates energy on \(\mathcal{O}_1\).

@@

Using \(x(t) := (q_u(t),p_u(t))\) as shorthand, let
\(x(0) \in \mathcal{O}_1\) be a nonzero initial condition of the acrobot.
Suppose the orbit \(x(\R)\) is confined within 
\(\bar{\mathcal{O}}_1\), and does not exit through \(E_\pi\) into the rotation
zone.
Since \(\bar{\mathcal{O}}_1\) is compact, the Birkhoff Theorem \cite{birkhoff}
implies:
\begin{itemize}
    \item The positive limit set \(L_+\) of \(x(t)\) is non-empty, compact, and
        invariant.
    \item The solution \(x(t)\) asymptotically tends to \(L_+\).
\end{itemize}
Since the acrobot gains energy on \(\mathcal{O}_1\), the positive limit set of
\(x(t)\) must be the largest invariant subset of \(E_\pi\). 
From our discussion on pseudo-polar coordinates (in particular the extension
~\eqref{eqn:alpha-dot-boundary} of \(\dot{\alpha}\) to \(\bar{\mathcal{O}}_1\)),
we know that \(\dot{\alpha} \geq 0\) on \(E_\pi\), with equality if and only if 
\(\alpha \in \{0,\pi\}\).
There are two possibilities: either \(E_\pi\) itself is invariant, or the
largest invariant subset of \(E_\pi\) is \(\{(\pi,0)\}\).
To rule out the first possibility, take the derivative of \(E(q_u,p_u)\) at the
\(p_u\)-axis to get
\[
    \dot{E}(0,p_u) = -\frac{g}{5l} \sin(q_a)p_u
    .
\]
This is non-zero everywhere except at the origin, which means it is non-zero on
\(E_\pi\). 
This means \(E_\pi\) is not invariant, so the positive limit set of \(x(t)\)
must be the set \(\{(\pi,0)\}\).
Hence, \(x(t)\) converges to \((\pi,0)\), which means \(x(0) \in \Pi^+\).
Since we know that \(\Pi^+\) is a set of measure zero in \(\SxR\), almost every
other orbit initialized in \(\mathcal{O}_1\) must escape through the boundary
\(E_\pi\) into the rotation domain.

@@

Suppose we initialize the acrobot at \((\beta,\rho) = (0,\rho_0)\).
One full rotation amounts to \(\beta\) traversing \(2\pi\) rad in a clockwise
direction; that is, when \(\beta\) goes from \(0\) to \(\sign{\rho_0}2\pi\).
Since \(\dot{\beta}\) does not change sign on \(\mathcal{R}\),
there is a well-defined Poincar\'{e} map describing how the pseudo-radius
\(\rho\) changes each time the orbit \((q(t),p(t))\) hits the \(p_u\) axis, \ie,
every time \(\beta\) changes by \(2\pi\).
We define this Poincar\'{e} map as
\(P_\mathcal{R}(\rho_0) := \hat{\rho}(\sign{\rho_0}2\pi,\rho_0,I)\),
which expands into
\[
    P_\mathcal{R}(\rho_0) = \rho_0 + I \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0)
    + R(\sign{\rho_0}2\pi,\rho_0,I)
    .
\]

@@

By the same arguments presented in this section, the Poincar\'{e} section
satisfies \(\abs{P_\mathcal{R}(\rho_0)} \leq \abs{\rho_0} - \gamma\) 
when \(I < 0\).

@@

\section{Experimental Results}
% TODO: This might be a separate chapter

>>>>

\section{The Acrobot Constraint}
One may be tempted to design a constraint of the form 
\(q_a = \bar{q}_a\sin(\theta)\),
with \(\theta := \arctan_2(p_u,q_u)\), since a similar approach 
was so effective for the VLP in Chapter \ref{ch:vlp}.
Unfortunately, this constraint is not regular, and it is difficult to find any
VNHC of the form \(q_a = f(\theta)\) where regularity can be proven easily.
Instead, we will develop a constraint \(h(q,p) = q_a - f(p_u)\) because these
constraints are always regular (as per Proposition \ref{prop:acrobot-fpu-regular}).

@@

By the same arguments presented in this section, the Poincar\'{e} section
satisfies
\[
    P_\mathcal{O}(\mu_0) \leq \mu_0 - \gamma
    ,
\] 
when \(I < 0\). 
In this case, the constraint dissipates energy on \(\mathcal{O}_1\).

@@

Using \(x(t) := (q_u(t),p_u(t))\) as shorthand, let
\(x(0) \in \mathcal{O}_1\) be a nonzero initial condition of the acrobot.
Suppose the orbit \(x(\R)\) is confined within 
\(\bar{\mathcal{O}}_1\), and does not exit through \(E_\pi\) into the rotation
zone.
Since \(\bar{\mathcal{O}}_1\) is compact, the Birkhoff Theorem \cite{birkhoff}
implies:
\begin{itemize}
    \item The positive limit set \(L_+\) of \(x(t)\) is non-empty, compact, and
        invariant.
    \item The solution \(x(t)\) asymptotically tends to \(L_+\).
\end{itemize}
Since the acrobot gains energy on \(\mathcal{O}_1\), the positive limit set of
\(x(t)\) must be the largest invariant subset of \(E_\pi\). 
From our discussion on pseudo-polar coordinates (in particular the extension
~\eqref{eqn:alpha-dot-boundary} of \(\dot{\alpha}\) to \(\bar{\mathcal{O}}_1\)),
we know that \(\dot{\alpha} \geq 0\) on \(E_\pi\), with equality if and only if 
\(\alpha \in \{0,\pi\}\).
There are two possibilities: either \(E_\pi\) itself is invariant, or the
largest invariant subset of \(E_\pi\) is \(\{(\pi,0)\}\).
To rule out the first possibility, take the derivative of \(E(q_u,p_u)\) at the
\(p_u\)-axis to get
\[
    \dot{E}(0,p_u) = -\frac{g}{5l} \sin(q_a)p_u
    .
\]
This is non-zero everywhere except at the origin, which means it is non-zero on
\(E_\pi\). 
This means \(E_\pi\) is not invariant, so the positive limit set of \(x(t)\)
must be the set \(\{(\pi,0)\}\).
Hence, \(x(t)\) converges to \((\pi,0)\), which means \(x(0) \in \Pi^+\).
Since we know that \(\Pi^+\) is a set of measure zero in \(\SxR\), almost every
orbit initialized in \(\mathcal{O}_1\) must escape through the boundary
\(E_\pi\) into the rotation domain.

Escaping once into the rotation domain does not guarantee the acrobot
orbit will not return to \(\mathcal{O}_1\).
Indeed, the orbit could escape through \(E_\pi\) and return to \(\mathcal{O}_1\)
several times.
However, the orbit must eventually remain outside the closure of 
\(\mathcal{O}_1\). 
If not, the positive limit set would once again be
\(E_\pi\); by the same argument as above, the orbit must converge to 
\((\pi,0)\).

We conclude that all orbits beginning in \(\mathcal{O}_1\) will, in finite time,
escape the closure of \(\mathcal{O}_1\) and remain in the rotation domain
forever after.

@@

Suppose we initialize the acrobot at \((\beta,\rho) = (0,\rho_0)\).
One full rotation amounts to \(\beta\) traversing \(2\pi\) rad in a clockwise
direction; that is, when \(\beta\) goes from \(0\) to \(\sign{\rho_0}2\pi\).
Since \(\dot{\beta}\) does not change sign on \(\mathcal{R}\),
there is a well-defined Poincar\'{e} map describing how the pseudo-radius
\(\rho\) changes each time the orbit \((q(t),p(t))\) hits the \(p_u\) axis, \ie,
every time \(\beta\) changes by \(2\pi\).
We define this Poincar\'{e} map as
\[
    P_\mathcal{R}(\rho_0) := \hat{\rho}\left(\sign{\rho_0}2\pi,\rho_0,I\right)
    ,
\]
which expands into
\[
    P_\mathcal{R}(\rho_0) = \rho_0 + I \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0)
    + R(\sign{\rho_0}2\pi,\rho_0,I)
    .
\]

@@

By the same arguments presented in this section, the Poincar\'{e} section
satisfies 
\[
    \abs{P_\mathcal{R}(\rho_0)} \leq \abs{\rho_0} - \gamma
    ,
\]
when \(I < 0\).

@@

\section{Experimental Results}
In this section, we test our VNHC on the physical acrobot built by
\citet{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
Since it is not a simple acrobot, this experiment will test whether VNHCs are
robust to model uncertainties.
As we will see, our VNHC can inject energy on the general acrobot model.

\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{images/xingbo_acrobot.png}
    \caption{The acrobot built by Wang \cite{xingbo_thesis}.}
    \label{fig:xingbo-acrobot}
\end{figure}

Wang's acrobot  has inertia matrix as in ~\eqref{eqn:general-acrobot-inertia}
and potential function as in ~\eqref{eqn:general-acrobot-potential}.
The physical parameters are written explicitly in Table
\ref{tab:acrobot-parameters}.

\begin{table}
    \centering
    \caption{Physical parameters for the real acrobot, as measured by
    \citet{xingbo_thesis}.}
    \label{tab:acrobot-parameters}
    \begin{tabular}{ccccccccc}
        \toprule
        $m_t$ & $m_l$ & $d_t$ & $d_l$ & $l_t$ & $l_l$ & $J_t$ & $J_l$ & $g$ \\
        \midrule
        0.2112 & 0.1979 & % mass
        0.148 & 0.145 & % total length
        0.073 & 0.083 & % COM distance
        0.00129 & 0.00075 & % moment of inertia
        9.81 \\ % gravity
        \bottomrule
    \end{tabular}
\end{table}

This acrobot's actuator is a servo motor with a built-in PID controller, so we
can set \(q_a = \arctan\left(I p_u\right)\) directly without needing to solve
for the VNHC's torque input \(\tau\).
The control parameter \(I\) needs to be small, but it also needs to be large
enough to overcome friction in the servo motor.
Since the acrobot's mass and length values are small, its momentum is
also small.
Through various experiments, we determined that \(I = 10\) is sufficient to
ensure the actuator can fully rotate in its allowable range
\(q_a \in \left[ -\frac{\pi}{2}, \frac{\pi}{2}\right]\).

\subsection{Simulations}

To compute the energy of the nominal pendulum, we evaluate the system at the
VNHC \(q_a = 0\), which yields
\[
    E(q_u,p_u) \approx 396.5501 p_u^2 + 0.5997(1 - \cos(q_u))
    .
\]
Hence, the level set of \(E_\pi\) of the energy \(E(\pi,0)\) is parameterized by
\begin{align*}
    p_u &\approx \pm \sqrt{\frac{1.1994}{793.1001}(2 - (1 - \cos(q_u))}
    ,
\end{align*}
for each \(q_u \in ]-\pi,\pi]\).

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.8\linewidth]{images/acrobot_orbit.png}
    \caption{A simulation of the acrobot from \cite{xingbo_thesis}.}
    \label{fig:acrobot-orbit}
\end{figure}

We initialize the simulation for this acrobot at 
\((q_u,p_u) = \left(\frac{\pi}{32},0 \right)\) and plot the resulting orbit in 
Figure \ref{fig:acrobot-orbit}.
The nominal energy level set \(E_\pi\) is outlined in black. 
Because this acrobot is not simple, \(E_\pi\) is not the boundary between
rotations and oscillations.
The oscillation domain is actually much larger: orbits rotate once they hit the
\(p_u\)-axis at \(\abs{p_u} \approx 0.15\), while \(E_\pi\) intersects the
\(p_u\)-axis at \(\abs{p_u} \approx 0.055\).
Despite this, we see that eventually the acrobot remains outside \(E_\pi\).
The points where the orbit exits \(E_\pi\) are highlighted with black stars,
with the final departure highlighted in red.
At this final departure, the acrobot begins rotating and continues to gain
energy over time.

To verify that the acrobot will consistently begin rotating, we use a Monte
Carlo method \cite{montecarlo} to simulate this
acrobot \(1000\) times.
At each iteration, we initialize the acrobot randomly inside the sublevel set
\[
    \left\{(q_u,p_u) \in \SxR \mid
    E(q_u,p_u) \leq E\left(\frac{\pi}{32},0\right)\right\}
    ,
\] 
and measure how long it takes the acrobot to begin rotating.
The results are plotted in Figure \ref{fig:acrobot-mc}.
The acrobot always rotates within 10--35 seconds.

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{images/acrobot_mc.png}
    \caption{Monte Carlo simulation results for the acrobot from
    \cite{xingbo_thesis}.}
    \label{fig:acrobot-mc}
\end{figure}

\subsection{Physical Experiments}
To complete this chapter, we perform a four experiments with the physical
acrobot:
\begin{enumerate}
    \item An undisturbed test to validate that the acrobot reaches rotations.
    \item A test where we stop the acrobot once it has started rotating.
    \item A test where we push the acrobot in its direction of motion after it has
        started rotating.
    \item A test where we push the acrobot against its direction of motion.
\end{enumerate}

The acrobot has an encoder at the pivot which measures \(q_u\) and
\(\dot{q}_u\), and the servo provides a measurement of \(q_a\).
We estimate \(\dot{q}_a\) through sequential values of \(q_a\) and
compute \(p_u\) through \(p_u = e_1^T M(q) \dot{q}\).
The actuator value at iteration \(k \in \mathbb{Z}_{> 0}\) is assigned through
the equation
\[
    q_a^{k} = \arctan(I p_u^{k-1})
    .
\]

\subsubsection*{Test 1: No Disturbances}
For this test we initialize the acrobot at 
\((q_u,p_u) \approx \left(\frac{\pi}{8},0\right)\). 
The resulting orbit is shown in Figure \ref{fig:acrobot-unperturbed-orbit},
which is clearly gaining energy over time.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\linewidth]{images/acrobot_unperturbed_orbit.png}
    \caption{The orbit of the physical acrobot during an unperturbed test.}
    \label{fig:acrobot-unperturbed-orbit}
\end{figure}

\subsubsection*{Test 2: Stopping the Acrobot}
For this test we initialize the acrobot at 
\((q_u,p_u) \approx \left(\pi,0\right)\) and let it run for 15
seconds, then stop the acrobot as it reaches the bottom of its arc.
The resulting orbit is shown in Figure \ref{fig:acrobot-stopped-orbit}.
The blue curves correspond to the orbit before the disturbance, while the red
spiral shows that the acrobot continues to oscillate after it is stopped.
Despite the disturbance, it gains energy and eventually reaches a rotation
again.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\linewidth]{images/acrobot_stopped_orbit.png}
    \caption{The orbit of the physical acrobot before (blue) and after (red) it
    is stopped.}
    \label{fig:acrobot-stopped-orbit}
\end{figure}

\subsubsection*{Test 3: Pushing the Acrobot Forwards}
To see how the acrobot responds when pushed in its direction of motion, we allow
the acrobot to rotate freely for 15 seconds and then give it a push in its
direction of motion.
The orbit in Figure \ref{fig:acrobot-fpush-orbit} shows that the acrobot speeds
up to rotate with energy \(E(0,0.22)\), but then slows down until it reaches a
stable rotation with energy \(E(0,-0.195)\).

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{images/acrobot_fpush_orbit.png}
    \caption{The orbit of the physical acrobot before (blue) and after (red) it
    is pushed in its current direction of motion.}
    \label{fig:acrobot-fpush-orbit}
\end{figure}

\subsubsection*{Test 4: Pushing the Acrobot Backwards}
In this final experiment, we test whether the acrobot can easily change
directions when pushed against its current direction of motion.
We again allow the acrobot to rotate freely for 15 seconds, then push it the
opposite way.
The orbit in Figure \ref{fig:acrobot-rpush-orbit} demonstrates that the acrobot
responds by readily changing direction, and quickly achieves its maximum speed
with energy \(E(0,0.195)\).

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{images/acrobot_rpush_orbit.png}
    \caption{The orbit of the physical acrobot before (blue) and after (red) it
    is pushed against its current direction of motion.}
    \label{fig:acrobot-rpush-orbit}
\end{figure}

\subsection{Summary of Experimental Results}
The simulations and experiments in this section show that VNHCs are excellent
tools for injecting energy even in non-simple acrobots. 
Furthermore, the experimental results show that the energy gain is robust
against a variety of disturbances.
Finally, the two push tests suggest that Wang's acrobot constrained by our VNHC
will gain energy on \(\mathcal{O}_2(0.195)\).
